---
title: "Predicting Sediment Budgets"
author: "Matthew J Cashman"
date: "October 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = getwd())
```

#Import and Prep Data
```{r Import and merge off previous script, message=FALSE, warning=FALSE}

root.dir = getwd()
source(paste0(root.dir,"/import_merge_gis_field.R"))
p_load(caret, e1071, corrplot,ggbiplot)
```

#Explore Bed Predictions off XS data
##Bed - Variable removal and MV- BoxCox
```{r Bed Predictions combined, message=FALSE, warning=FALSE, include=FALSE}
#Choose Bed
data <- XS_ws_site %>%
  filter(Class=="Bed")
data_num <- data[,-1:-3]

apply(data_num, 2, skewness)

data_num <- data_num[,-nearZeroVar(data_num)]
data_num <- data_num %>%
  select(-fill.rate)
data_num <- data_num %>%
   select(-p_dev_high, -p_dev_med, -p_dev_lo, -p_dev_os,
          -S25RANG, -area_imper, -area_canopy, -G25RANG, -S100RANG,
          -range_slopes, -grad_range_100m, -p_ag_sum,
          -G25STD, -S25STD, -p_forest_sum,-S25MAX,-S100MAX,-max_slope,-G25MAX, -grad_max_100m,
         -G25MEAN, -p_dev_sum, -grad_mean_100m, -min_slope, -grad_min_100m, -std_slope, -grad_std_100m,
         -shreve, -strahler, -p_herb, -p_shrub, -S100STD, -p_hay, -pcent_canopy, -p_decid_forest)

trans <- preProcess(as.data.frame(data_num), method = c("BoxCox"))
data_trans <- predict(trans, as.data.frame(data_num))
```
```{r Transform predictor and dependent}
ggdensity(data_trans$net.rate, 
          main = "Density plot")
ggqqplot(data_trans$net.rate)
shapiro.test((data_trans$net.rate))

p_load(rcompanion)

depvar.transformed<-transformTukey(data_trans$net.rate+1,    plotit=TRUE)

ggdensity(depvar.transformed, 
          main = "Density plot")
ggqqplot(depvar.transformed)
shapiro.test((depvar.transformed))
data_trans$net.rate <- depvar.transformed

```

##Bed - Check collinearity

```{r Bed Check Collinearity}
correlations <- cor(data_trans)

as.data.frame(correlations) %>%
  dplyr::select(net.rate) %>%
  rownames_to_column() %>%
  arrange(desc(abs(net.rate)))

corrplot(correlations, order="hclust", method="pie")

p_load(PerformanceAnalytics)

# my_data <- data_trans[, c(1,3,4,5,6,7)]
# chart.Correlation(my_data, histogram=TRUE, pch=19)

highCorr <- findCorrelation(correlations, cutoff = 0.75)
colnames(data_trans[,highCorr])
```

##Bed - Bed PCA
```{r Bed PCA}
ir.pca <- prcomp(data_trans,
                center = TRUE,
                scale. = TRUE) 
plot(ir.pca, type = "l")
summary(ir.pca)
```

##Bed - Bed PCA2
```{r Bed PCA Pt2}
g <- ggbiplot(ir.pca, obs.scale = 1, var.scale = 1, groups = data$Class, ellipse = TRUE, circle = TRUE)
print(g)
```

##Bed - LM predictions pre and post-transformation
```{r LM predictions post-transform}
fit1 <- lm(data=data_trans, net.rate~.)
summary(fit1)
step1 <- step(fit1)
summary(step1)
plot(step1)
```


#Explore Bank Predictions off Bank Pin Data
##Bank - Variable removal and MV-Boxcox
```{r Bank variable removal and transformation}
data <- Bank_ws_site %>%
  select(-Avg_NormDiff_cm_per_day) %>%
  na.omit() ##5 sites are being omitted here because they did not match between XS and bank locations!!!

data_num <- data[,-1]

#apply(data_num, 2, skewness)

data_num <- data_num[,-nearZeroVar(data_num)]

data_num <- data_num %>%
   select(-p_dev_high, -p_dev_med, -p_dev_lo, -p_dev_os,
          -S25RANG, -area_imper, -area_canopy, -G25RANG, -S100RANG,
          -range_slopes, -grad_range_100m, -p_ag_sum,
          -G25STD, -S25STD, -p_forest_sum,-S25MAX,-S100MAX,-max_slope,-G25MAX, -grad_max_100m,
         -G25MEAN, -p_dev_sum, -grad_mean_100m, -min_slope, -grad_min_100m, -std_slope, -grad_std_100m,
         -shreve, -strahler, -p_herb, -p_shrub, -S100STD, -p_hay, -pcent_canopy, -p_decid_forest)

trans <- preProcess(as.data.frame(data_num), method = c("BoxCox"))
data_trans <- predict(trans, as.data.frame(data_num))

ggdensity(data_trans$Avg_NormCumDiff_cm_per_day, 
          main = "Density plot")
ggqqplot(data_trans$Avg_NormCumDiff_cm_per_day)
shapiro.test((data_trans$Avg_NormCumDiff_cm_per_day)) #No transformation necessary

```
##Bank - Check collinearity
```{r Check Bank Collinearity}
p_load(corrplot)

data_trans1 <- data_trans
correlations <- cor(data_trans1)

as.data.frame(correlations) %>%
  dplyr::select(Avg_NormCumDiff_cm_per_day) %>%
  rownames_to_column() %>%
  arrange(desc(abs(Avg_NormCumDiff_cm_per_day)))

corrplot(correlations, order="AOE")

highCorr <- findCorrelation(correlations, cutoff = 0.75)
colnames(data_trans1[,highCorr])

```
##Bank - PCA
```{r Bank PCA}
ir.pca <- prcomp(data_trans,
                 center = TRUE,
                 scale. = TRUE) 
plot(ir.pca, type = "l")
g <- ggbiplot(ir.pca, obs.scale = 1, var.scale = 1, ellipse = TRUE, circle = TRUE)
print(g)
```

##Bank - LM predictions pre and post-transformation
```{r Bank Predictions}
ggplot(data = data_trans, aes(x=p_bare, y=log(Avg_NormCumDiff_cm_per_day+1)))+
  geom_point()

#LM

fit <- lm(data=data_trans, Avg_NormCumDiff_cm_per_day~LAT+LONG+Bank_Angle+bankheight+channelwidth+floodwidth+channelflood_ratio+contr_area+pcent_imp+mean_slope)
summary(fit)
step <- step(fit)
summary(step)

fit <- lm(data=data_trans, Avg_NormCumDiff_cm_per_day~LAT+LONG+strahler+shreve+Bank_Angle+
            bankheight+channelwidth+floodwidth+channelflood_ratio+contr_area+
            pcent_canopy+pcent_imp+max_slope+S100MEAN+p_bare+p_hay+p_crops+p_dev_sum)
summary(fit)
step <- step(fit)
summary(step)
plot(step)
```

#Explore Floodplain Predictions off Tile Data
```{r Floodplain Predictions combined}
data <- tile_ws_site[,-1] %>%
  select(-NameofRiver, -Thickness_mm, -diff_time)
str(data)
data_num <- data

apply(data_num, 2, skewness)

data_num <- data_num[,-nearZeroVar(data_num)]

data_num <- data_num %>%
   select(-X.y, -p_dev_high, -p_dev_med, -p_dev_lo, -p_dev_os,
          -S25RANG, -area_imper, -area_canopy, -G25RANG, -S100RANG,
          -range_slopes, -grad_range_100m, -p_ag_sum,
          -G25STD, -S25STD, -p_forest_sum,-S25MAX,-S100MAX,-max_slope,-G25MAX, -grad_max_100m,
         -G25MEAN, -p_dev_sum, -grad_mean_100m, -min_slope, -grad_min_100m, -std_slope, -grad_std_100m,
         -shreve, -strahler, -p_herb, -p_shrub, -S100STD, -p_hay, -pcent_canopy, -p_decid_forest)

trans <- preProcess(as.data.frame(data_num), method = c("BoxCox"))
data_trans <- predict(trans, as.data.frame(data_num))

ggplot(data=data, aes(y=deposit_mm_yr, x=log(elev_above_thalweg*Dist_from_thalweg)))+
  geom_point()+
  facet_wrap(~Site, scales = "free_x")+
  geom_smooth(method="lm")

#
fit <- lm(data=data_num[,-1:-2], log(deposit_mm_yr+1)~.)
summary (fit)
step <- step(fit)
summary(step)

#Site tile data averaged in deposition by site
fit <- lm(data=data_num[,-1:-2], log(deposit_mm_yr+1)~.)
summary (fit)
step <- step(fit)
summary(step)
```

